---
title: "p-values in Pathology Articles"
subtitle: "A Text Analysis of PubMed Articles Using Regular Expression"
author: "Serdar Balcı, MD, Pathologist"
date: '2019-01-07'
bibliography: crossref.bib
output: 
  html_notebook: 
    code_folding: hide
    fig_caption: yes
    highlight: tango
    theme: paper
    toc_depth: 5
  html_document: 
    code_folding: hide
    df_print: kable
    fig_caption: yes
    highlight: pygments
    keep_md: yes
    theme: paper
    toc_depth: 5
---

# p-values in Pathology Articles {.tabset .tabset-fade .tabset-pills}


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'figure/', echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, eval = TRUE, tidy = TRUE, comment = NA)
```


```{r library, include=FALSE}
library(tidyverse)
```


---

## Introduction

---

### Background

#### p-values

@Nead_2018 


#### significance term throughout years


```{r}
library(europepmc)

significanTerm <- europepmc::epmc_hits_trend(query = "significan*", period = 1980:2017)

significanTerm

```


```{r}
library(ggplot2)
ggplot(significanTerm, aes(year, query_hits / all_hits)) + 
  geom_point() + 
  geom_line() +
  xlab("Year published") + 
  ylab("Proportion of 'significan*' \n articles in Europe PMC")
```



### Aim



## Analysis

---

### Materials and Methods



#### Regular Expression

```
{regular expression code}
p[<>=]0\.\d{1,7}|p.[<>=].0\.\d{1,7}|significant|nonsignificant|non-significant
```


---

#### Data retriveal from PubMed using EDirect

- Articles are downloaded as `xml`.



```{r Search PubMed write ModPathol2016plus data as xml, eval=FALSE, include=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
    myTerm,
    "esearch -db pubmed -query \"'Mod Pathol'[Journal] \" -datetype PDAT -mindate 2016 -maxdate 3000 | efetch -format xml > data/ModPathol2016plus.xml \n"
)
Sys.sleep(1)
repeat {
    Sys.sleep(0.1)
    if (rstudioapi::terminalBusy(myTerm) == FALSE) {
        print("Code Executed")
        break
    }
}
```


- extract abstracts from `xml`.


```{r extract journal names from 2018 xml, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
myTerm,
"xtract -input data/Turkey_2018.xml -pattern PubmedArticle -element MedlineCitation/PMID PubDate/Year Journal/ISSN ISOAbbreviation > data/Turkey2018.csv \n"
)
Sys.sleep(1)
repeat {
Sys.sleep(0.1)
if (rstudioapi::terminalBusy(myTerm) == FALSE) {
print("Code Executed")
break
}
}
```






----


```
-element MedlineTA  NLMCatalogRecord/NlmUniqueID -def 'na' -sep '\t' -block TitleAlternate/Title   -element TitleAlternate/Title

"xtract -input data/nlmcatalog_result_currentlyindexed.xml -pattern NCBICatalogRecord -element ISSNLinking -def 'na' -sep ' ' -block TitleAlternate/Title -if TitleAlternate/Title@Sort -equals N -element TitleAlternate/Title  > data/nlmcatalog.csv \n"

-sep '\t'


NLMCatalogRecord/NlmUniqueID ISSNLinking
```


---


```{r nlmcatalog, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
myTerm,
"xtract -input data/nlmcatalog_result_currentlyindexed.xml -pattern NCBICatalogRecord -tab '|' -element NLMCatalogRecord/NlmUniqueID -block ISSNLinking -tab '|' -element ISSNLinking -block Title -if Title@Sort -equals N -def 'na' -tab '|' -element TitleAlternate/Title > data/nlmcatalog.csv \n"
)
Sys.sleep(1)
repeat {
Sys.sleep(0.1)
if (rstudioapi::terminalBusy(myTerm) == FALSE) {
print("Code Executed")
break
}
}

``` 


```{r eval=FALSE, include=FALSE}
library(readr)
nlmcatalog <- read_delim("data/nlmcatalog.csv", 
                         delim = "|",
                         escape_double = FALSE,
                         col_names = FALSE, 
                         trim_ws = TRUE)

```

```{r eval=FALSE, include=FALSE}

library(xml2)
data <- read_xml("data/nlmcatalog_result_currentlyindexed.xml")

# Point locations
point <- data %>% xml_find_all("//pointer")
point %>% xml_attr("latitude") %>% as.numeric()
point %>% xml_attr("longitude") %>% as.numeric()

# Start time
data %>% 
  xml_find_all("//start-valid-time") %>% 
  xml_text()

# Temperature
data %>% 
  xml_find_all("//temperature[@type='hourly']/value") %>% 
  xml_text() %>% 
  as.integer()



```




---

## Results

---



---

## Discussion

---


### Comment

---



---


## Feedback

---

[Serdar Balcı, MD, Pathologist](https://github.com/sbalci) would like to hear your feedback: https://goo.gl/forms/YjGZ5DHgtPlR1RnB3

This document will be continiously updated and the last update was on `r Sys.Date()`.



---

<div id="disqus_thread"></div>
<script>

/**
*  RECOMMENDED CONFIGURATION VARIABLES: EDIT AND UNCOMMENT THE SECTION BELOW TO INSERT DYNAMIC VALUES FROM YOUR PLATFORM OR CMS.
*  LEARN WHY DEFINING THESE VARIABLES IS IMPORTANT: https://disqus.com/admin/universalcode/#configuration-variables*/
/*
var disqus_config = function () {
this.page.url = PAGE_URL;  // Replace PAGE_URL with your page's canonical URL variable
this.page.identifier = PAGE_IDENTIFIER; // Replace PAGE_IDENTIFIER with your page's unique identifier variable
};
*/
(function() { // DON'T EDIT BELOW THIS LINE
var d = document, s = d.createElement('script');
s.src = 'https://https-sbalci-github-io.disqus.com/embed.js';
s.setAttribute('data-timestamp', +new Date());
(d.head || d.body).appendChild(s);
})();
</script>
<noscript>Please enable JavaScript to view the <a href="https://disqus.com/?ref_noscript">comments powered by Disqus.</a></noscript>



---

## Back to Main Page for Bibliographic Analysis

---

[Main Page for Bibliographic Analysis](https://sbalci.github.io/pubmed/BibliographicStudies.html)


---

## References  

---

https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#inline_references

The Use of “Trend” Statements to Describe Statistically Nonsignificant Results in the Oncology Literature  
JAMA Oncol. 2018;4(12):1778-1779. doi:10.1001/jamaoncol.2018.4524  
https://jamanetwork.com/journals/jamaoncology/fullarticle/2705969  


---



```{r knitcitations}
# knitcitations::citep("10.1186/s12905-016-0336-0")
```





---

