---
title: "Pancreas Studies with Various Bibliometric Tools"
subtitle: "Pancreas Studies in Pathology Journals"
author: "Serdar Balcı, MD, Pathologist"
date: '`r format(Sys.Date())`'
output: 
  html_notebook: 
    code_folding: hide
    fig_caption: yes
    highlight: kate
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 5
    toc_float: yes
bibliography: bib/PancreasStudies.bib
---

<!-- Open all links in new tab-->  
<!-- <base target="_blank"/>    -->


<!-- Go to www.addthis.com/dashboard to customize your tools --> <script type="text/javascript" src="//s7.addthis.com/js/300/addthis_widget.js#pubid=ra-5bc36900a405090b">  
</script> 


```{r setup global_options chunk settings, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	error = FALSE,
	fig.height = 8,
	fig.width = 12,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	comment = NA,
	tidy = TRUE, 
	eval = TRUE
# fig.path = 'figures/',

)
```

```{r load library, include=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(foreign))
suppressMessages(library(tidylog))
suppressMessages(library(janitor))
suppressMessages(library(googledrive))
```



# Introduction

- Selected Pathology Journals are used for the query^[[Selected Pathology Journals](PancreasStudies/data/journal_list_from_my_collections)]




# PubMed

---

## Collection with Article Features


```{r}
PBPathJWArticleFeatures <- xlsx::read.xlsx("/Users/serdarbalciold/JournalWatchPBPath/DataForCloud/PBPathArticles.xlsx", sheetIndex = 1)
```





```{r eval=FALSE, include=FALSE}
# library(googledrive)
# googledrive::drive_auth()
# gdrive_folder <-  "https://drive.google.com/drive/folders/1QIzvkJOK6STkfAyb8rI36J34J7kGQpwT"
# 
# https://drive.google.com/drive/folders/1QIzvkJOK6STkfAyb8rI36J34J7kGQpwT
# 
# drive_find(pattern="PBPathJWArticles/PBPathArticles.xlsx")

# gdrls <- googledrive::drive_ls(gdrive_folder)

# dplyr::pull(gdrls, id) %>% 
    # purrr::walk(~googledrive::drive_download(googledrive::as_id(.x), path = "/data")


# googledrive::drive_download(file = "~/PBPathJWArticles/PBPathArticles.xlsx",
#                                 # path = "data",
#                                 overwrite = TRUE)
```



## My NCBI » Collections > PBPathJournalWatch  

https://www.ncbi.nlm.nih.gov/sites/myncbi/collections/58240634


```{r}
myNCBICollectionsPBPathJournalWatch <- read.csv(file = "https://www.ncbi.nlm.nih.gov/portal/utils/streamer.cgi?request=%3CRequest%20version%3D%271%27%3E%3CCommand%20name%3D%27collection-save%27%3E%09%3CParam%20name%3D%27id%27%3E58240634%3C%2FParam%3E%09%3CParam%20name%3D%27format%27%3Ecsv%3C%2FParam%3E%09%3CParam%20name%3D%27sortKey%27%3Edate%3C%2FParam%3E%09%3CParam%20name%3D%27sortOrder%27%3Edescending%3C%2FParam%3E%20%20%3C%2FCommand%3E%3C%2FRequest%3E")
```


```{r}
# text_myNCBICollectionsPBPathJournalWatch <- readr::read_lines_raw( "https://www.ncbi.nlm.nih.gov/portal/utils/streamer.cgi?request=%3CRequest%20version%3D%271%27%3E%3CCommand%20name%3D%27collection-save%27%3E%09%3CParam%20name%3D%27id%27%3E58240634%3C%2FParam%3E%09%3CParam%20name%3D%27format%27%3Etxt%3C%2FParam%3E%09%3CParam%20name%3D%27sortKey%27%3Edate%3C%2FParam%3E%09%3CParam%20name%3D%27sortOrder%27%3Edescending%3C%2FParam%3E%20%20%3C%2FCommand%3E%3C%2FRequest%3E")
```


---

## peeriodicals

https://peeriodicals.com/peeriodicals/pancreatobiliary-pathology-society

---

## Data Download from PubMed via EDirect

- Data is downloaded from PubMed via Edirect^[https://dataguide.nlm.nih.gov/edirect/overview.html]

--- 


## rentrez

---

# europepmc

---

# WOS

[WOS.Rmd](WOS.Rmd)

```{r}
PMID_List_for_WOS_pancreas <- paste0("PMID=(", pmid_only_path_journals$PMID, ") OR")

write(PMID_List_for_WOS_pancreas,
      here::here("PancreasStudies/data/PMID_List_for_WOS_pancreas.txt")
)

```







# Scopus

[scopus.Rmd](scopus.Rmd)


# Vosviewer

Author Thesaurus is used.^[[Author Thesaurus](data/thesaurus_authors_myupdates.txt)]



Code for interactive graph.^[[Interactive Graph Code](data/VOSviewer_interactive_graphcode.txt)]






# Cite Space

# CitNetExplorer

# Lens

```
(citation_id_type: "pmid"  AND citation_id: 28984336)  OR (citation_id_type: "pmid"  AND citation_id: 21682559)
```


```{r}
pmid_only_path_journals <- 
readr::read_delim("PancreasStudies/data/pmid_only_path_journals.txt", "\n", col_names = "PMID", col_types = "c")


pmid_only_path_journals2 <- paste0("citation_id:(",pmid_only_path_journals$PMID, ") OR ", collapse = "")

write(pmid_only_path_journals2,
"PancreasStudies/data/lensSearch.txt"
)
```




# openknowledgemaps

Open Knowledge Maps (2019). Overview of research on intraductal tubulopapillary neoplasm. Retrieved from https://openknowledgemaps.org/map/cc04c53301920a93d451aa2ce3010243/ [30 Mar 2019]. 


# bibliometrix

# ORCID

# Semantic Scholar

[SemanticScholar.Rmd](SemanticScholar.Rmd)


# Microsoft Academic

# Scopus

# Altmetrics



# Grid

https://grid.ac/


# ROR

https://ror.org/search?page=1&query=turkey


# Cobalmetrics

https://cobaltmetrics.com/

- Toward privacy-preserving altmetrics exploration with Cobaltmetrics and ORCID

https://medium.com/thunken/toward-privacy-preserving-altmetrics-exploration-with-cobaltmetrics-and-orcid-f5ab6fa7898a


# OpenCitations

http://opencitations.net

http://opencitations.net/download

http://opencitations.net/corpus

https://github.com/opencitations

- The OpenCitations Data Model  
https://figshare.com/articles/Metadata_for_the_OpenCitations_Corpus/3443876

- Creating Open Citation Data with BCite  
https://semsci.github.io/SemSci2018/papers/1/bcite-semsci2018.html

https://github.com/opencitations/bcite


# Open Knowledge Maps

Open Knowledge Maps (2019). Overview of research on (pathology[affiliation] or patoloji[affiliation]) and (turkey[affiliation] or türkiye[affiliation]). Retrieved from https://openknowledgemaps.org/map/e4dbf6c89015e402d6d7d560ab79e17a/ [30 Mar 2019]. 


https://openknowledgemaps.org/index



# iris.ai

https://the.iris.ai/map/02f89d24-0415-43ab-878a-853ec17c5a59

# lens.org


<iframe src="https://player.vimeo.com/video/295099740" width="640" height="360" frameborder="0" webkitallowfullscreen mozallowfullscreen allowfullscreen></iframe>
<p><a href="https://vimeo.com/295099740">The Lens</a> from <a href="https://vimeo.com/user10734565">Cambia</a> on <a href="https://vimeo.com">Vimeo</a>.</p>

<iframe src="https://player.vimeo.com/video/245867897" width="640" height="360" frameborder="0" allowfullscreen></iframe>

<iframe src="https://player.vimeo.com/video/246041050" width="640" height="360" frameborder="0" allowfullscreen></iframe>

https://www.lens.org/lens/scholar/search/results?dateFilterField=year_published&filterMap=%7B%7D&orcids=0000-0002-7852-3851&orderBy=%2Bpublished&preview=true&previewType=SCHOLAR_ANALYSIS&regexEnabled=false

https://www.lens.org/lens/scholar/search/results?q=(author.affiliation.name:%20%22pathology%22%20%20OR%20author.affiliation.name:%20%22patoloji%22)%20%20AND%20(author.affiliation.name:%20%22Turkey%22%20%20OR%20author.affiliation.name:%20%22T%C3%BCrkiye%22)

https://www.lens.org/lens/scholar/search/analysis?q=(author.affiliation.name:%20%22neurosurgery%22%20%20OR%20author.affiliation.name:%20%22sport%22%20)%20AND%20(abstract:%20%22sport%22%20%20OR%20title:%20%22sport%22)&page=0&limit=10&orderBy=%2Bscore&dateFilterField=year_published&preview=false&regexEnabled=false



# contentmine

http://contentmine.org/

https://github.com/ContentMine







# To Do

- find a way to auto import journal names
- remove last OR
- Auto run posiible analysis via cron
- Find a way to auto copy into googledrive
- push to blog, netlify ...
- Add from
-- [Sources.Rmd](Sources.Rmd)  
-- [Bibliyometri.Rmd](Bibliyometri.Rmd)  




# About & Feedback


[Serdar Balcı, MD, Pathologist](https://github.com/sbalci) would like to hear your feedback: https://goo.gl/forms/YjGZ5DHgtPlR1RnB3

This document will be continiously updated and the last update was on `r Sys.time()`.

Serdar Balci, MD, Pathologist  
drserdarbalci@gmail.com  
https://rpubs.com/sbalci/CV   

---

<script id="dsq-count-scr" src="//https-sbalci-github-io.disqus.com/count.js" async></script>

<div id="disqus_thread"></div>
<script>

/**
*  RECOMMENDED CONFIGURATION VARIABLES: EDIT AND UNCOMMENT THE SECTION BELOW TO INSERT DYNAMIC VALUES FROM YOUR PLATFORM OR CMS.
*  LEARN WHY DEFINING THESE VARIABLES IS IMPORTANT: https://disqus.com/admin/universalcode/#configuration-variables*/
/*
var disqus_config = function () {
this.page.url = PAGE_URL;  // Replace PAGE_URL with your page's canonical URL variable
this.page.identifier = PAGE_IDENTIFIER; // Replace PAGE_IDENTIFIER with your page's unique identifier variable
};
*/
(function() { // DON'T EDIT BELOW THIS LINE
var d = document, s = d.createElement('script');
s.src = 'https://https-sbalci-github-io.disqus.com/embed.js';
s.setAttribute('data-timestamp', +new Date());
(d.head || d.body).appendChild(s);
})();
</script>
<noscript>Please enable JavaScript to view the <a href="https://disqus.com/?ref_noscript">comments powered by Disqus.</a></noscript>

---

# Libraries Used

```{r}
citation()
```

```
citation("tidyverse")
citation("foreign")
citation("tidylog")
citation("janitor")
citation("jmv")
citation("tangram")
citation("finalfit")
citation("summarytools")
citation("ggstatplot")
citation("readxl")
```


```{r eval=FALSE, include=FALSE}
citation("tidyverse")
citation("foreign")
citation("tidylog")
citation("janitor")
citation("jmv")
citation("tangram")
citation("finalfit")
citation("summarytools")
citation("ggstatplot")
citation("readxl")
```

```{r}
sessionInfo()
```


---

# References




