---
title: "Pancreas Studies in Various Bibliometric Tools"
subtitle: "Pancreas Studies in Pathology Journals"
author: "Serdar Balcı, MD, Pathologist"
date: '`r format(Sys.Date())`'
output: 
  html_notebook: 
    code_folding: hide
    fig_caption: yes
    highlight: monochrome
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 5
    toc_float: yes
---

<!-- Open all links in new tab-->  
<!-- <base target="_blank"/>    -->


<!-- Go to www.addthis.com/dashboard to customize your tools --> <script type="text/javascript" src="//s7.addthis.com/js/300/addthis_widget.js#pubid=ra-5bc36900a405090b">  
</script> 


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, 
                      # fig.path = 'figures/',
                      echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, eval = TRUE, tidy = TRUE, comment = NA)
```

```{r library, include=FALSE}
library(tidyverse)
```




# Introduction

- Selected Pathology Journals are used for the query^[[Selected Pathology Journals](PancreasStudies/data/journal_list_from_my_collections)]




# PubMed

## Data Download from PubMed via EDirect

- Data is downloaded from PubMed via Edirect^[https://dataguide.nlm.nih.gov/edirect/overview.html]

- 


## rentrez

# europepmc



# WOS

[WOS.Rmd](WOS.Rmd)

```{r}
PMID_List_for_WOS_pancreas <- paste0("PMID=(", pmid_only_path_journals$PMID, ") OR")

write(PMID_List_for_WOS_pancreas,
      here::here("PancreasStudies/data/PMID_List_for_WOS_pancreas.txt")
)

```







# Scopus

[scopus.Rmd](scopus.Rmd)


# Vosviewer

Author Thesaurus is used.^[[Author Thesaurus](data/thesaurus_authors.txt)]

Code for interactive graph.^[[Interactive Graph Code](data/VOSviewer_interactive_graphcode.txt)]






# Cite Space

# CitNetExplorer

# Lens

```
(citation_id_type: "pmid"  AND citation_id: 28984336)  OR (citation_id_type: "pmid"  AND citation_id: 21682559)
```

```{r}
pmid_only_path_journals <- 
read_delim("PancreasStudies/data/pmid_only_path_journals.txt", "\n", col_names = "PMID", col_types = "c")

pmid_only_path_journals2 <- paste0("(citation_id_type: 'pmid'  AND citation_id:",pmid_only_path_journals$PMID, ") OR ", collapse = "")

write(pmid_only_path_journals2,
"PancreasStudies/data/lensSearch.txt"
)
```




# openknowledgemaps

Open Knowledge Maps (2019). Overview of research on intraductal tubulopapillary neoplasm. Retrieved from https://openknowledgemaps.org/map/cc04c53301920a93d451aa2ce3010243/ [30 Mar 2019]. 


# bibliometrix

# ORCID

# Semantic Scholar

[SemanticScholar.Rmd](SemanticScholar.Rmd)


# Microsoft Academic

# Scopus

# Altmetrics



# Grid

https://grid.ac/


# ROR

https://ror.org/search?page=1&query=turkey


# Cobalmetrics

https://cobaltmetrics.com/

- Toward privacy-preserving altmetrics exploration with Cobaltmetrics and ORCID

https://medium.com/thunken/toward-privacy-preserving-altmetrics-exploration-with-cobaltmetrics-and-orcid-f5ab6fa7898a


# OpenCitations

http://opencitations.net

http://opencitations.net/download

http://opencitations.net/corpus

https://github.com/opencitations

- The OpenCitations Data Model  
https://figshare.com/articles/Metadata_for_the_OpenCitations_Corpus/3443876

- Creating Open Citation Data with BCite  
https://semsci.github.io/SemSci2018/papers/1/bcite-semsci2018.html

https://github.com/opencitations/bcite


# Open Knowledge Maps

Open Knowledge Maps (2019). Overview of research on (pathology[affiliation] or patoloji[affiliation]) and (turkey[affiliation] or türkiye[affiliation]). Retrieved from https://openknowledgemaps.org/map/e4dbf6c89015e402d6d7d560ab79e17a/ [30 Mar 2019]. 


https://openknowledgemaps.org/index



# iris.ai

https://the.iris.ai/map/02f89d24-0415-43ab-878a-853ec17c5a59

# lens.org

https://www.lens.org/lens/scholar/search/results?dateFilterField=year_published&filterMap=%7B%7D&orcids=0000-0002-7852-3851&orderBy=%2Bpublished&preview=true&previewType=SCHOLAR_ANALYSIS&regexEnabled=false

https://www.lens.org/lens/scholar/search/results?q=(author.affiliation.name:%20%22pathology%22%20%20OR%20author.affiliation.name:%20%22patoloji%22)%20%20AND%20(author.affiliation.name:%20%22Turkey%22%20%20OR%20author.affiliation.name:%20%22T%C3%BCrkiye%22)

https://www.lens.org/lens/scholar/search/analysis?q=(author.affiliation.name:%20%22neurosurgery%22%20%20OR%20author.affiliation.name:%20%22sport%22%20)%20AND%20(abstract:%20%22sport%22%20%20OR%20title:%20%22sport%22)&page=0&limit=10&orderBy=%2Bscore&dateFilterField=year_published&preview=false&regexEnabled=false


# To Do

- find a way to auto import journal names
- remove last OR
- Auto run posiible analysis via cron
- Find a way to auto copy into googledrive
- push to blog, netlify ...
- Add from
-- [Sources.Rmd](Sources.Rmd)  
-- [Bibliyometri.Rmd](Bibliyometri.Rmd)  




# About & Feedback


[Serdar Balcı, MD, Pathologist](https://github.com/sbalci) would like to hear your feedback: https://goo.gl/forms/YjGZ5DHgtPlR1RnB3

This document will be continiously updated and the last update was on `r Sys.Date()`.

---

<script id="dsq-count-scr" src="//https-sbalci-github-io.disqus.com/count.js" async></script>

<div id="disqus_thread"></div>
<script>

/**
*  RECOMMENDED CONFIGURATION VARIABLES: EDIT AND UNCOMMENT THE SECTION BELOW TO INSERT DYNAMIC VALUES FROM YOUR PLATFORM OR CMS.
*  LEARN WHY DEFINING THESE VARIABLES IS IMPORTANT: https://disqus.com/admin/universalcode/#configuration-variables*/
/*
var disqus_config = function () {
this.page.url = PAGE_URL;  // Replace PAGE_URL with your page's canonical URL variable
this.page.identifier = PAGE_IDENTIFIER; // Replace PAGE_IDENTIFIER with your page's unique identifier variable
};
*/
(function() { // DON'T EDIT BELOW THIS LINE
var d = document, s = d.createElement('script');
s.src = 'https://https-sbalci-github-io.disqus.com/embed.js';
s.setAttribute('data-timestamp', +new Date());
(d.head || d.body).appendChild(s);
})();
</script>
<noscript>Please enable JavaScript to view the <a href="https://disqus.com/?ref_noscript">comments powered by Disqus.</a></noscript>

---

# References




