---
title: "Pathology Articles From Turkey"
description: |
  Collections and Peeriodicals of selected Pathology Articles From Turkey
author:
  - name: Serdar Balcı 
    url: https://www.serdarbalci.com/
    affiliation: MD Pathologist
    affiliation_url: https://sbalci.github.io/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r library, include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(markdown))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(tidyRSS))
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



<!-- Open all links in new tab-->  
<!-- <base target="_blank"/>   -->

---

# Pathology Articles From Turkey Peeriodical

https://peeriodicals.com/peeriodicals/pathology-articles-from-turkey


---

# RSS Feed From PubMed

```
(pathology[Affiliation] OR patoloji[Affiliation]) AND (Turk*[Affiliation] OR Türk*[Affiliation]) 
```

`r icon::fa("rss")`
[Click Here for the RSS Link](https://eutils.ncbi.nlm.nih.gov/entrez/eutils/erss.cgi?rss_guid=10UQVR81oEgBsJdu9HkJM34nE3gKkc0-VJoAwxWLhi1d2qRs7u)



```{r}
# https://cran.r-project.org/web/packages/tidyRSS/vignettes/tidyrss.html
PAFT_rss1 <- tidyRSS::tidyfeed("https://eutils.ncbi.nlm.nih.gov/entrez/eutils/erss.cgi?rss_guid=10UQVR81oEgBsJdu9HkJM34nE3gKkc0-VJoAwxWLhi1d2qRs7u") 
```



```{r}
PAFT_rss2 <- feedeR::feed.extract("https://eutils.ncbi.nlm.nih.gov/entrez/eutils/erss.cgi?rss_guid=10UQVR81oEgBsJdu9HkJM34nE3gKkc0-VJoAwxWLhi1d2qRs7u")
```






---

# PubMed Collection


My NCBI » Collections > Pathology Articles From Turkey

`(pathology[Affiliation] OR patoloji[Affiliation]) AND (Turk*[Affiliation] OR Türk*[Affiliation])` AND `WOS Journals with Pathology Topic`

https://www.ncbi.nlm.nih.gov/sites/myncbi/collections/58299810/

<a href="https://www.ncbi.nlm.nih.gov/sites/myncbi/serdar.balci.1/collections/58299810/public/">View my collection, "Pathology Articles From Turkey" from NCBI</a>




```{r download PAFT data, include=FALSE}
PAFT <- readr::read_csv("https://www.ncbi.nlm.nih.gov/sites/myncbi/collections/58299810/?title=Open%20as%20spreadsheet&")

```


```{r Prepare Table, eval=FALSE, include=FALSE}
PAFT2 <- PAFT %>% 
  mutate(Link = paste0("https://www.ncbi.nlm.nih.gov", URL)) %>% 
  select(Title, Link) %>% 
  head(n=25)
```


**Style1**

```{r Print Table, eval=FALSE, include=FALSE, layout="l-body-outset"}
rmarkdown::paged_table(PAFT2)
```

**Style2**

```{r Print Table 2, eval=FALSE, include=FALSE, layout="l-body-outset"}
PAFT2 %>% 
  gt::gt()
```

**Style3**

```{r Print Table 3, eval=FALSE, include=FALSE, layout="l-body-outset"}
PAFT2 %>% 
  print.data.frame()
```

**Style4**

```{r Print Table 4, eval=FALSE, include=FALSE, layout="l-screen-inset"}
PAFT2 %>% 
  knitr::kable()
```



```{r Print Table Molecular, eval=FALSE, include=FALSE, layout="l-screen-inset"}

# PancreasPBPathArticles$Title <- kableExtra::cell_spec(
#   PancreasPBPathArticles$Title,
#   popover = kableExtra::spec_popover(
#     content = PancreasPBPathArticles$Journal
#   )
# )


PAFT %>% 
  filter(Molecular == TRUE) %>%
  select(Title, PubMedLink) %>% 
  kableExtra::kable(escape = FALSE) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped",
                          "hover",
                          "condensed",
                          "responsive"),
    fixed_thead = TRUE,
    full_width = TRUE
  ) %>%
  kableExtra::scroll_box(height = "300px")
```




## Acknowledgments {.appendix}

## Footnotes {.appendix}

## References {.appendix}

## Document Information {.appendix}

Last updated on `r Sys.Date()`.




