---
title: "Bibliographic Studies"
author: "Serdar Balci, MD, Pathologist"
date: '`r format(Sys.time())`'
output: 
  html_notebook: 
    code_folding: none
    fig_caption: yes
    highlight: kate
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_float: yes
---

**"MeSH Terms and Author Key Words In Pathology Articles From Turkey"**



```{r}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, tidy = TRUE, error = TRUE)
```

```{r}
library(rmarkdown)
```


# Background

PubMed collection from [National Library of Medicine](https://www.ncbi.nlm.nih.gov/pubmed/), has the most comprehensive information about peer reviewed articles in medicine.

[MeSH Terms](https://www.nlm.nih.gov/pubs/factsheets/mesh.html) is a controlled vocabulary that is used to label PubMed articles according to their content. It is done by experts in National Library of Medicine. Keywords are lables that are given by authors of the article. Both are included in a PubMed record of an article.

[EDirect](https://dataguide.nlm.nih.gov/edirect/install.html) is a valuable tool to identify various features of published articles that are indexed in PubMed.

# Aim

In this analysis we aimed to identify the most common topics investigated in Turkey and published in pathology journals. This may reveal the common research topics Turkish pathologists are interested. We extracted most common MeSH terms and keywords from PubMed articles using [EDirect](https://dataguide.nlm.nih.gov/edirect/overview.html):
[MeSH Terms Pathology Articles From Turkey](https://sbalci.github.io/pubmed/MeSH_Terms_Pathology_Articles_From_Turkey.html)


# Materials and Methods

**Methods:**

Packages used for analysis are given. `Tidyverse` is used for data manipulation, and [rstudioapi](https://github.com/rstudio/rstudio/issues/2193) to run e-utilities commands from RStudio.


```{r load -if not present install- required packages}
usePackage <- function(p)
{
if (!is.element(p, installed.packages()[, 1]))
install.packages(p, dep = TRUE)
require(p, character.only = TRUE)
}

usePackage("tidyverse")
usePackage("rstudioapi")
usePackage("pander")
```


Pathology Journal ISSN List was retrieved from [In Cites Clarivate](https://jcr.incites.thomsonreuters.com/), and Journal Data Filtered as follows: 

`JCR Year: 2016 Selected Editions: SCIE,SSCI Selected Categories: 'PATHOLOGY' Selected Category Scheme: WoS`


```{r Get ISSN List from data downloaded from WoS}
# # Get ISSN List from data downloaded from WoS
# ISSNList <- JournalHomeGrid <- read_csv("/Users/serdarbalciold/RepTemplates/pubmed/data/JournalHomeGrid.csv", skip = 1) %>%
# select(ISSN) %>%
# filter(!is.na(ISSN)) %>%
# t() %>%
# paste("OR ", collapse = "") # add OR between ISSN List
# 
# ISSNList <- gsub(" OR $", "" , ISSNList) # to remove last OR
```


Data is retrieved from PubMed via [e-Utilities](https://dataguide.nlm.nih.gov/).

The search formula for PubMed is generated as "ISSN List AND Country[Affiliation]" as described in [advanced search of PubMed](https://www.ncbi.nlm.nih.gov/pubmed/advanced).

```{r Generate Search Formula For Pathology Journals AND Countries}
# # Generate Search Formula For Pathology Journals AND Countries
# searchformula <-
# paste("'", ISSNList, "'", " AND ", "Turkey[Affiliation]")
# write(searchformula, "/Users/serdarbalciold/RepTemplates/pubmed/data/searchformula.txt")
```

Articles are downloaded as `xml`.

```{r Search PubMed, write data as xml}
# myTerm <- rstudioapi::terminalCreate(show = FALSE)
# rstudioapi::terminalSend(
#     myTerm,
#     "esearch -db pubmed -query \"$(cat /Users/serdarbalciold/RepTemplates/pubmed/data/searchformula.txt)\" -datetype PDAT -mindate 1900 -maxdate 3000 | efetch -format xml > /Users/serdarbalciold/RepTemplates/pubmed/data/PathologyTurkey.xml \n"
# )
# Sys.sleep(1)
# repeat {
#     Sys.sleep(0.1)
#     if (rstudioapi::terminalBusy(myTerm) == FALSE) {
#         print("Code Executed")
#         break
#     }
# }
```


MeSH terms are extracted from xml. [Common terms](https://www.nlm.nih.gov/bsd/indexing/training/CHK_010.html) are excluded and [major topics](https://www.nlm.nih.gov/bsd/disted/meshtutorial/principlesofmedlinesubjectindexing/majortopics/) are selected.

```{r extract major MeSH topics -excluding common tags- from xml, message=FALSE, warning=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
myTerm,
"xtract -input /Users/serdarbalciold/RepTemplates/pubmed/data/PathologyTurkey.xml -pattern MeshHeading -if DescriptorName@MajorTopicYN -equals Y -or QualifierName@MajorTopicYN -equals Y -element DescriptorName| grep -vxf /Users/serdarbalciold/RepTemplates/pubmed/data/checktags.txt | sort-uniq-count-rank > /Users/serdarbalciold/RepTemplates/pubmed/data/PathologyTurkeyMeSH.csv \n"
)
Sys.sleep(1)
repeat {
Sys.sleep(0.1)
if (rstudioapi::terminalBusy(myTerm) == FALSE) {
print("Code Executed")
break
}
}
```

Keywords are extracted from `xml`.

```{r extract author keywords from xml, message=FALSE, warning=FALSE, results='asis'}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
myTerm,
"xtract -input /Users/serdarbalciold/RepTemplates/pubmed/data/PathologyTurkey.xml -pattern Keyword -element Keyword | sort-uniq-count-rank > /Users/serdarbalciold/RepTemplates/pubmed/data/authorkeywords.csv \n"
)
Sys.sleep(1)
repeat {
Sys.sleep(0.1)
if (rstudioapi::terminalBusy(myTerm) == FALSE) {
print("Code Executed")
break
}
}
```


The retrieved information was compiled in a table.

```{r message=FALSE, warning=FALSE}
library(readr)

authorkeywords <- read_table2("/Users/serdarbalciold/RepTemplates/pubmed/data/authorkeywords.csv", col_names = c("frequency", "author key word"), cols(
  frequency = col_integer(),
  `author key word` = col_character()
), guess_max = 100)

authorkeywords <- authorkeywords %>% 
    head(n=20)

PathologyTurkeyMeSH <- read_table2("/Users/serdarbalciold/RepTemplates/pubmed/data/PathologyTurkeyMeSH.csv", col_names = c("frequency", "MeSH term"), cols(
  frequency = col_integer(),
  `MeSH term` = col_character()
), guess_max = 100)

PathologyTurkeyMeSH <-PathologyTurkeyMeSH %>% 
    head(n = 20)
```

# Results

_**Most common 20 author supplied keywords are given below.**_

```{r results = 'asis'}
print(authorkeywords)
# pander::pander(authorkeywords, justify = "left", caption = "Most common 20 author supplied keywords")
```

_**Most common 20 MeSH terms are given below**_

```{r results = 'asis'}

# pander::pander(PathologyTurkeyMeSH, justify = "left", caption = "Most common 20 MeSH terms")

print(PathologyTurkeyMeSH)

```



# Comments

**Comment:**
We may conclude that adenocarcinomas, breast and thyroid tumors, and immunohistochemistry are the most common research topics. Lung, kidney and ovary are also common research topics.


**Future Work:**
* The changes in frequencies of topics throughout time
* The distribution of topics among institutions




---


# Feedback

[Serdar Balcı, MD, Pathologist](https://github.com/sbalci) would like to hear your feedback: https://goo.gl/forms/YjGZ5DHgtPlR1RnB3

This document will be continiously updated and the last update was on `r Sys.time()`.

---

# Back to Main Menu

[Main Page for Bibliographic Analysis](https://sbalci.github.io/pubmed/BibliographicStudies.html)