---
title: "Retracted Pathology Articles in PubMed"
output: html_notebook
---

# Aim

To evaluate the retracted articles in PubMed.



# Materials and Methods

Data downloaded on 20.10.2019.

[currentlyindexed_nlmcatalog_result.xml](https://www.ncbi.nlm.nih.gov/portal/utils/file_backend.cgi?Db=nlmcatalog&HistoryId=NCID_1_12059143_130.14.18.48_5555_1571596396_818866540_0MetA0_S_HStore&QueryKey=1&Sort=PubDate&Filter=all&CompleteResultCount=5244&Mode=file&View=xml&p$l=Email&portalSnapshot=%2Fprojects%2Fentrez%2Fpubmed%2FPubMedGroup@1.146&BaseUrl=&PortName=live&RootTag=NLMCatalogRecordSet&DocType=NLMCatalogRecordSet%20PUBLIC%20%22-%2F%2FNLM%2F%2FDTD%20NLMCatalogRecordSet,%201st%20June%202017%2F%2FEN%22%20%22https://www.nlm.nih.gov/databases/dtd/nlmcatalogrecordset_170601.dtd%22&FileName=&ContentType=xml)


[Retracted_Publication_sb_OR_Retraction_of_Publication_sb_pubmed_result.xml](https://www.ncbi.nlm.nih.gov/portal/utils/file_backend.cgi?Db=pubmed&HistoryId=NCID_1_12059143_130.14.18.48_5555_1571596396_818866540_0MetA0_S_HStore&QueryKey=18&Sort=&Filter=all&CompleteResultCount=14043&Mode=file&View=xml&p$l=Email&portalSnapshot=%2Fprojects%2Fentrez%2Fpubmed%2FPubMedGroup@1.146&BaseUrl=&PortName=live&RootTag=PubmedArticleSet&DocType=PubmedArticleSet%20PUBLIC%20%22-%2F%2FNLM%2F%2FDTD%20PubMedArticle,%201st%20January%202019%2F%2FEN%22%20%22https://dtd.nlm.nih.gov/ncbi/pubmed/out/pubmed_190101.dtd%22&FileName=&ContentType=xml)


[Retracted_Publication_sb_OR_Retraction_of_Publication_sb_timeline.csv](https://www.ncbi.nlm.nih.gov/pubmed?p$l=Email&Mode=download&term=Retracted%20Publication[sb]%20OR%20Retraction%20of%20Publication[sb]&dlid=timeline&filename=timeline.csv&bbid=&p$debugoutput=off)



- Find Retractions Using PubMed and My NCBI!

https://hslnews.wordpress.com/2013/05/14/retractions/

- Comment Correction Type

https://www.ncbi.nlm.nih.gov/books/NBK3827/#pubmedhelp.Comment_Correction_Type



> Comment Correction Type
> 
> The data in these fields are citations to other associated journal publications, e.g., comments or errata. Often these link to the respective citation. Comments/Corrections data can be retrieved by the search term that follows each type:
> 
>     Comment in: hascommentin
>     Comment on: hascommenton
>     Corrected and republished in: hascorrectedrepublishedin
>     Corrected and republished from: hascorrectedrepublishedfrom
>     Dataset use reported in: hasassociatedpublication
>     Dataset described in: hasassociateddataset
>     Erratum in: haserratumin
>     Erratum for: haserratumfor
>     Expression of concern in: hasexpressionofconcernin
>     Expression of concern for: hasexpressionofconcernfor
>     Original Report in: hasoriginalreportin
>     Republished in: hasrepublishedin
>     Republished from: hasrepublishedfrom
>     Retracted and republished in: hasretractedandrepublishedin
>     Retracted and republished from: hasretractedandrepublishedfrom
>     Retraction in: hasretractionin
>     Retraction of: hasretractionof
>     Summary for patients in: hassummaryforpatientsin
>     Update in: hasupdatein
>     Update of: hasupdateof
> 

- Errata, Retractions, and Other Linked Citations in PubMed

https://www.nlm.nih.gov/bsd/policy/errata.html


<!-- Not used -->
<!-- ## Data retriveal from PubMed using EDirect  -->

<!-- Article PMID downloaded as `txt`. -->

```{r Search PubMed download PMID, eval=FALSE, include=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
    myTerm,
    "esearch -db pubmed -query \"(pathology[Affiliation] OR patoloji[Affiliation]) AND (Turkey[Affiliation] OR Türkiye[Affiliation])\" -datetype PDAT -mindate 1800 -maxdate 3000 | \ efetch -format uid > data/pubmed_result_TurkPath_uid.txt \n"
)
Sys.sleep(1)
repeat {
    Sys.sleep(0.1)
    if (rstudioapi::terminalBusy(myTerm) == FALSE) {
        print("Code Executed")
        break
    }
}
```



```{r Search local PubMed with downloaded PMID extract data as xml, eval=FALSE, include=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
    myTerm,
    "cat data/pubmed_result_TurkPath_uid.txt | \ fetch-pubmed -path /Volumes/Agu2018/PubMed > data/pubmed_result_TurkPath.xml \n"
)
Sys.sleep(1)
repeat {
    Sys.sleep(0.1)
    if (rstudioapi::terminalBusy(myTerm) == FALSE) {
        print("Code Executed")
        break
    }
}
```



# Analysis


A trend graph of retracted articles derived from Europe PMC #RStats #europepmc {europepmc} 📦 #evergreenreviewgraph #research #bibliography #bibliometrics



```{r europepmc}
library(europepmc)
```


```{r europepmc search}
retractedArticlesPerTotalArticles <- 
    europepmc::epmc_hits_trend(
    query = "(PUB_TYPE:'Retracted Publication' OR PUB_TYPE:'Retraction of Publication')",
    period = 1980:2017
    )

retractedArticlesPerTotalArticles$Affiliation <- "All"

retractedArticlesPerTotalArticles_Turkey <- 
    europepmc::epmc_hits_trend(
    query = "(AFF:'Turkey') AND (PUB_TYPE:'Retracted Publication' OR PUB_TYPE:'Retraction of Publication')",
    period = 1980:2017
    )

retractedArticlesPerTotalArticles_Turkey$Affiliation <- "Turkey"

retractedArticles <- dplyr::bind_rows(
retractedArticlesPerTotalArticles,
retractedArticlesPerTotalArticles_Turkey)

```



```{r graph europepmc search}
library(ggplot2)
ggplot(retractedArticles, 
       aes(year, 
           1000*(query_hits / all_hits),
           color = Affiliation)
       ) + 
  geom_point() + 
  geom_line() +
  xlab("Year published") + 
  ylab("Proportion of Retracted Articles \n data: Europe PMC, 20.10.2019 \n 1000x")
```












```{r extract year journal name from xml, message=FALSE, warning=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
myTerm,
"xtract -input data/RetractedPubMed/Retracted_Publication_sb_OR_Retraction_of_Publication_sb_pubmed_result.xml -pattern PubmedArticle -tab \"|\" -sep \";\" -def \"NA\" -element MedlineCitation/PMID ArticleTitle Journal/ISSN ISOAbbreviation PubDate/Year > data/RetractedPubMed/pubmed_result_retracted.csv \n"
)
Sys.sleep(1)
repeat {
Sys.sleep(0.1)
if (rstudioapi::terminalBusy(myTerm) == FALSE) {
print("Code Executed")
break
}
}
```



```{r read extracted data}
pubmed_result_retracted <-
    readr::read_delim(
        file = here::here("data",
                          "RetractedPubMed",
                          "pubmed_result_retracted.csv"),
        delim = "|",
        escape_double = FALSE,
        col_names = FALSE,
        trim_ws = TRUE
    )

pubmed_result_retracted <- 
pubmed_result_retracted %>%
    select(
        PMID = X1,
        title = X2,
        journalISSN = X3, 
        journalName = X4,
        year = X5
    )
```






```{r extract journal name and topics from xml, message=FALSE, warning=FALSE}
myTerm <- rstudioapi::terminalCreate(show = FALSE)
rstudioapi::terminalSend(
myTerm,
"xtract -input data/currentlyindexed_nlmcatalog_result.xml -pattern NCBICatalogRecord -tab \"|\" -sep \";\" -def \"NA\" -element MedlineTA TitleAlternate BroadJournalHeadingList MeshHeading/DescriptorName ISSN > data/RetractedPubMed/journal_properties.csv \n"
)
Sys.sleep(1)
repeat {
Sys.sleep(0.1)
if (rstudioapi::terminalBusy(myTerm) == FALSE) {
print("Code Executed")
break
}
}
```


```{r read clean journal properties}
journal_properties <- readr::read_delim(here::here("data", "RetractedPubMed" , "journal_properties.csv"), 
    "|", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE)

journal_properties <- journal_properties %>% 
    select(journal = X1,
           topic = X4,
           ISSN = X5)



journal_properties <- journal_properties %>% 
    mutate(journalName = str_split(journal, ";")) %>%
  unnest(cols = c(journalName)) %>% 
    mutate(journalTopic = str_split(topic, ";")) %>%
  unnest(cols = c(journalTopic)) %>% 
    mutate(journalISSN = str_split(ISSN, ";")) %>%
  unnest(cols = c(journalISSN)) %>% 
    select(journalName,
           journalTopic,
           journalISSN) %>% 
    unique()

```



```{r combine retracted articles with their journal info}
pubmed_result_retracted <- pubmed_result_retracted %>% 
    left_join(journal_properties, by = c("journalISSN"))
```


```{r find the journal topics in which most retracted articles were}
commonTerms <- c("Medicine", "Science", "Research")

(pubmed_result_retracted %>% 
    select(journalTopic, year) %>% 
    filter(!journalTopic %in% commonTerms) %>% 
            filter(!is.na(journalTopic)) %>% 
    group_by(journalTopic) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    head(10) %>% 
    pull(journalTopic) -> mostRetractedTopicsOfJournal)
```




This trend graph shows the major topics of the journals that have the most retractions. Data derived from PubMed #RStats  #research #bibliography #bibliometrics


```{r major topics of the journals that have the most retractions}
pubmed_result_retracted %>% 
    filter(journalTopic %in% mostRetractedTopicsOfJournal) %>% 
    filter(year <= 2018 & year >= 1990) %>% 
    select(journalTopic, year) %>% 
    group_by(journalTopic, year) %>% 
    tally() %>% 
    ggplot( 
       aes(year, 
           n,
           color = journalTopic
           )
       ) + 
  geom_point() + 
  geom_line() +
  xlab("Year published") + 
  ylab("Proportion of Retracted Articles \n data: Europe PMC, 20.10.2019")
```









